---
title: "ReproducibleResearch_Assignment1"
author: "F.Lopez"
date: "April 5, 2016"
output: html_document
---

This document provides the answers to the Course Project 1, part of the Reproducible Research course.

List of questions to be answered in this assignment:

1. Code for reading in the dataset and/or processing the data
2. Histogram of the total number of steps taken each day
3. Mean and median number of steps taken each day
4. Time series plot of the average number of steps taken
5. The 5-minute interval that, on average, contains the maximum number of steps
6. Code to describe and show a strategy for imputing missing data
7. Histogram of the total number of steps taken each day after missing values are imputed
8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

## 1 Code for reading in the dataset and/or processing the data
```{r, echo = TRUE}
d<- read.csv (file = "activity.csv", header = TRUE)
d$date<-as.Date(d$date,format='%Y-%m-%d')
```

## 2 Histogram of the total number of steps taken each day

```{r, echo=TRUE}
d1 <- aggregate(x = d[c("steps","interval")],
                     FUN = sum,
                     by = list(Group.date = d$date), na.rm=TRUE)

##  Histogram
hist(d1$steps,
     col = "lightgray",
     border = "black", 
     xlab = "Interval",
     main = "Histogram - Total Steps per day")
  abline(v = mean(d1$steps),
         col = "blue",
         lwd = 2)
  abline(v = median(d1$steps),
         col = "red",
         lwd = 2)
  legend(x = "topright", 
       c("Mean", "Median"),
       col = c("blue", "red"),
       lwd = c(2, 2, 2))
```


## 3. Mean and median number of steps taken each day

```{r, echo=TRUE}
d1mean <- setNames(aggregate(x = d[c("steps")],
                FUN = mean,
                by = list(d$date), na.rm=TRUE), c("Date", "Mean"))

d1median <- setNames(aggregate(x = d[c("steps")],
                    FUN = median,
                    by = list(d$date), na.rm=TRUE), c("Date", "Median"))

d1stats <- merge(d1mean,d1median,by="Date")

print(d1stats)
```


## 4. Time series plot of the average number of steps taken

```{r, echo=TRUE}
## aggregated table for graphic
d2mean <- setNames(aggregate(x = d[c("steps")],
                             FUN = mean,
                             by = list(d$interval), na.rm=TRUE), c("Interval", "MeanStep"))

## Interval with Max Average Steps
d2max<-subset(d2mean, d2mean$MeanStep==max(d2mean$MeanStep, na.rm = TRUE))

## plot histogram
plot(d2mean$Interval, d2mean$MeanStep, type="l", main = "Steps per Interval", xlab= "Interval", ylab= "Steps(Mean)", col="red" ,lwd=1)
    abline(v = d2max$Interval, col = "blue", lwd = 2)
    legend(x = "topright", 
           c("MaxSteps Interval"),
           col = c("blue"),
           lwd = c(2, 2, 2))
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps

```{r, echo=TRUE}
d2max<-subset(d2mean, d2mean$MeanStep==max(d2mean$MeanStep, na.rm = TRUE))
    print(d2max)
```


## 6. Code to describe and show a strategy for imputing missing data

NA values will be replaced with globla average

```{r, echo=TRUE}
## original data to be stored in a new variable called 'd3'
d3 <- d

##NA values in column 'Steps' are replaced with global average
d3$steps[which(is.na(d3$steps))] <- mean(d3$steps, na.rm = TRUE)

## example
head(d3)

```

## 7. Histogram of the total number of steps taken each day after missing values are imputed

```{r, echo=TRUE}
## aggregated data for histogram
d4 <- aggregate(x = d3[c("steps")],
                FUN = sum,
                by = list(Group.date = d$date), na.rm=TRUE)

## stats
d4mean <- setNames(aggregate(x = d4[c("steps")],
                             FUN = mean,
                             by = list( d4$Group.date), na.rm=TRUE), c("Date", "Mean"))

d4median <- setNames(aggregate(x = d4[c("steps")],
                               FUN = median,
                               by = list(d4$Group.date), na.rm=TRUE), c("Date", "Median"))

d4stats <- merge(d4mean,d4median,by="Date")

print(d4stats)

## histogram
hist(d4$steps,
       col = "lightgray",
       border = "black", 
       xlab = "SUM(steps)",
       main = "Histogram  Total Steps per day")
        abline(v = mean(d4$steps),
         col = "blue",
         lwd = 2)
        
        abline(v = median(d4$steps),
         col = "red",
         lty = 2,
         lwd = 2)
        
        legend(x = "topright", 
         c("Mean","Median"),
         col = c("blue","red"),
         lwd = c(2, 2, 2))
```

## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r, echo=TRUE}
d5<-d3
d5$day <- weekdays(d3$date)
d5$day <- ifelse(d5$day == "Saturday" | d5$day == "Sunday" ,"weekend", "weekday")
d5$day <- as.factor(d5$day)


## aggregated data for graphic
d5sum <- setNames(aggregate(x = d5[c("steps")],
                             FUN = sum,
                             by = list(d5$interval, d5$day), na.rm=TRUE), c("Interval", "day", "SumStep"))

## subsetting data for graphics
d5weekday<-subset(d5sum, d5sum$day=="weekday")
d5weekend<-subset(d5sum, d5sum$day=="weekend")

## plot
  par(mfrow=c(2,1))
        plot(d5weekday$Interval,d5weekday$SumStep, main="Interval vs SumStep [weekdays]", type = "l", col = "blue", xlab = "Interval", ylab = "SumStep")
        plot(d5weekend$Interval,d5weekend$SumStep, main="Interval vs SumStep [weekend]", type = "l", col = "red", xlab = "Interval", ylab = "SumStep")
```

## 9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report
See steps 1 to 8

-- end

